---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# knapsack

The goal of knapsack is to find the combination of objects with maximum value without exceeding a given weight. Three approaches are implemented, the brute force algorithm, greedy algorithm and dynamic algorithm. 

## Installation

You can install knapsack from github with:

```{r gh-installation, eval = FALSE}
#install.packages("devtools")
devtools::install_github("mariatreesa/RLab6")
```

## Examples

The knapsack objects used:

```{r echo=FALSE}
set.seed(42)
n = 1000000
knapsack_objects <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000)
  )

head(knapsack_objects)
```


The following code shows how to invoke the three algorithms for knapsack problem.

###Brute Force Algorithm

```{r example, echo=FALSE}
## brute force example
brute_force_knapsack <- function(x,W){
  # stop for erroneous input
  if(is.numeric(W)== F || is.data.frame(x) ==F){
    stop("Please enter valid inputs")
  }
  else if(W <= 0){
    stop("Please enter weight larger than 0")
  }


  # empty list
  value_knap <- list(value = c(0), elements = c())

  #number of items
  n <- nrow(x)

  # list of all possible combinations
  combs <- list()
  # get all possible combinations
  # We want as many items as possible in the bag
  for(i in seq_len(n)){

  combs[i] <- list(combn(n, i))

  }

  for(i in seq_len(n)){
    k = ncol(combs[[i]])

    # for each selection get the total value
    for (j in seq_len(k)) {
      # if sum of comb value is less than W store it in sel
      if((sum(x[combs[[i]][,j],1]) <= W) && sum(x[combs[[i]][,j],2]) > value_knap$value) {

        # max value from comb
        value_knap$value <- sum(x[combs[[i]][,j],2])


        # elements with maximum value and least weight
        value_knap$elements <- combs[[i]][,j]
      }
    }


  }

return(value_knap)

}

```

```{r}
brute_force_knapsack(x = knapsack_objects[1:8,],W=2000)
```


###Dynamic Algorithm

```{r example2, echo=FALSE}
## dynamic example
dynamic_knapsack <- function(x,W,fast = FALSE){

  if(is.numeric(W)== F || is.data.frame(x) ==F){
    stop("Please enter valid inputs")
  }
  else if(W <= 0){
    stop("Please enter weight larger than 0")
  }

  if(fast == FALSE){
   m <- matrix(0, ncol = (W+1), nrow = (nrow(x)+1))

   # The code below runs two for loops to get the maximum value
   #that can be extracted keeping the weight minimum

   for(i in 2:nrow(m)){
     for(j in 2:ncol(m)){
       if (x$w[i-1] > j-1){
        m[i, j] = m[i-1, j]}
       else if(j-x$w[i-1] >= 0){
          m[i, j] = max(m[i-1, j], m[i-1, ((j-x$w[i-1]))] + x$v[i-1])
        }
       else{
          m[i, j] = m[i-1, j]
        }

       }
   }
  }
  else{
    m = knapSackdynamic_cpp(W,x$w,x$v,nrow(x))
  }


   # the code below runs one for loop to get the elements that are used to get the value above
   val <- m[nrow(m),ncol(m)]
   elements <- c()

   for(row in nrow(m):2){
    if(!(val %in% m[row-1,])){
      elements <- c(elements, row-1)
      val = val - x$v[row-1]
    }

   }
   elements = sort(elements, decreasing = FALSE)

   result <- list("value" = round(m[nrow(m),ncol(m)]), "elements" = elements)
   return(result)

}

```

```{r}
dynamic_knapsack(x = knapsack_objects[1:8,],W=2000,fast = FALSE)
```


###Greedy Algorithm
```{r example3, echo=FALSE}
## greedy example

greedy_knapsack <- function(x,W, fast = FALSE ){

  if(is.numeric(W)== F || is.data.frame(x) ==F){
    stop("Please enter valid inputs")
  }
  else if(W <= 0){
    stop("Please enter weight larger than 0")
  }


  x$elements <- as.numeric(rownames(x))
  x$vw <- x$v/x$w
  x <- x[order(-x$vw),]
  x <- x[which(x$w <= W),]
  x$weight_sum <- cumsum(x$w)
  x <- x[which(x$weight_sum <= W),]
  if(fast == TRUE){
    knapsackvalue <- vectorSum(x$v)
  }else{
  knapsackvalue <- sum(x$v)
  }
  elements <- x$elements

   result <- list("value" = round(knapsackvalue), "elements" = elements)
   return(result)

}

```

```{r}
greedy_knapsack(x = knapsack_objects[1:8,],W=2000,fast = FALSE)
```


