---
title: "Knapsack Algorithm Implementation"
author: "Maria, Masinde, Omkar"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(knapsack)
```

The goal of knapsack package is to find the combination of objects with maximum value without exceeding a given weight. 

This package has the solutions for Knapsack problem using three algorithms.
1. Brute Force Algorithm
2. Dynamic Algorithm
3. Greedy Algorithm

The object used in knapsack_objects. This is a data frame with 2000 rows and 2 columns. The column data cointains the weight and value of objects.

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
set.seed(42)
n = 1000000
knapsack_objects <-
  data.frame(
    w=sample(1:4000, size = n, replace = TRUE),
    v=runif(n = n, 0, 10000)
  )
```


```{r}
head(knapsack_objects)
```



##The Usage of Package

###brute_force_knapsack(X,W)

This algorithm can be invoked as given below:

brute_force_knapsack(x = knapsak_objects[1:8,],W=2000)

Output

```{r}
brute_force_knapsack(x = knapsack_objects[1:8,],W=2000)

```

###dynamic_knapsack(x,W)

This algorithm can be invoked as given below:

dynamic_knapsack(x = knapsack_objects[1:8,],W=2000)

Output

```{r}
dynamic_knapsack(x = knapsack_objects[1:8,],W=2000)

```


###greedy_knapsack(x,W)

This algorithm can be invoked as given below:

greedy_knapsack(x = knapsack_objects[1:8,],W=2000)

Output

```{r}
greedy_knapsack(x = knapsack_objects[1:8,],W=2000)

```

##Speed of Algorithms

###brute_force_knapsack(x,W)

Time taken to run the algorithm for 16 objects

```{r}
starttime <- Sys.time()
brute_force_knapsack(x = knapsack_objects[1:16,],W=2000)
endtime <- Sys.time()

bruteforce_time = endtime - starttime
```

```{r echo=FALSE}
bruteforce_time
```


###dynamic_knapsack(x,W)

Time taken to run the algorithm for 500 objects

```{r}
starttime <- Sys.time()
dynamic_knapsack(x = knapsack_objects[1:500,],W=3500)
endtime <- Sys.time()

dynamic_time = endtime - starttime
```

```{r echo=FALSE}
dynamic_time
```


###dynamic_knapsack(x,W,fast=TRUE)

Time taken to run the algorithm for 500 objects using Rcpp

```{r}
starttime <- Sys.time()
dynamic_knapsack(x = knapsack_objects[1:500,],W=3500,fast = TRUE)
endtime <- Sys.time()

dynamic_time = endtime - starttime
```

```{r echo=FALSE}
dynamic_time
```

###greedy_knapsack(x,W)

Time taken to run the algorithm for 1000000 objects

```{r}
starttime <- Sys.time()
greedy_knapsack(x = knapsack_objects[1:1000000,],W=3500)
endtime <- Sys.time()

greedy_time = endtime - starttime
```

```{r echo=FALSE}
greedy_time
```

###greedy_knapsack(x,W,fast=TRUE) - usingRcpp

Time taken to run the algorithm for 1000000 objects using Rcpp

```{r}
starttime <- Sys.time()
greedy_knapsack(x = knapsack_objects[1:1000000,],W=3500, fast = TRUE)
endtime <- Sys.time()

greedy_time = endtime - starttime
```


```{r echo=FALSE}
greedy_time
```

We can clearly observe the advantage of using cpp function in above two examples. The code that used cpp in both Greedy Algorithm and Dynamic algorithm takes less time compared to the normal R algorithms.
