---
title: "Knapsack Algorithm Implementation"
author: "Sebastian Masinde Omkar"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(knapsack)
```

The goal of knapsack package is to find the combination of objects with maximum value without exceeding a given weight. 

This package has the solutions for Knapsack problem using three algorithms.
1. Brute Force Algorithm
2. Dynamic Algorithm
3. Greedy Algorithm

The object used in knapsack_objects. This is a data frame with 2000 rows and 2 columns. The column data cointains the weight and value of objects.

```{r}
head(knapsack_objects)
```



##The Usage of Package

###brute_force_knapsack(X,W)

This algorithm can be invoked as given below:

brute_force_knapsack(X = knapsak_objects[1:8],W=16)

Output

```{r}
brute_force_knapsack(X = knapsak_objects[1:8],W=2000)

```

###dynamic_knapsack(x,W)

This algorithm can be invoked as given below:

dynamic_knapsack(X = knapsak_objects[1:8],W=2000)

Output

```{r}
dynamic_knapsack(x = knapsak_objects[1:8],W=2000)

```


###greedy_knapsack(x,W)

This algorithm can be invoked as given below:

greedy_knapsack(X = knapsak_objects[1:8],W=2000)

Output

```{r}
greedy_knapsack(x = knapsak_objects[1:8],W=2000)

```

##Speed of Algorithms

###brute_force_knapsack(X,W)

Time taken to run the algorithm for 16 objects

```{r}
starttime <- system.time()
brute_force_knapsack(x = knapsak_objects[1:16],W=2000)
endtime <- system.time()

bruteforce_time = starttime - endtime
```

The time taken to run brute force algorithm for 16 objects is: 

```{r}
bruteforce_time
```


###dynamic_knapsack(X,W)

Time taken to run the algorithm for 500 objects

```{r}
starttime <- system.time()
dynamic_knapsack(x = knapsak_objects[1:500],W=3500)
endtime <- system.time()

dynamic_time = starttime - endtime
```

The time taken to run brute force algorithm for 500 objects is: 

```{r}
dynamic_time
```


###dynamic_knapsack(X,W)

Time taken to run the algorithm for 500 objects

```{r}
starttime <- system.time()
greedy_knapsack(x = knapsak_objects[1:1000000],W=3500)
endtime <- system.time()

greedy_time = starttime - endtime
```

The time taken to run brute force algorithm for 1000000 objects is: 

```{r}
greedy_time
```
